---
title: "merck_coding_interview"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

For a coding interview in R, it's crucial to be familiar with both base R and the tidyverse, as they offer different approaches to data manipulation and analysis. Here's a breakdown of the essentials for both:

### Base R Essentials:

1.  **Vectors**: Understanding how to create, subset, and manipulate vectors is fundamental. Know about numeric, character, logical, and factor vectors.

```{r}
# create 
data = c(1,2,3,5,6,2,3)

# basic slicing 
data_sub = data[1:4]

# numeric vector 
class(data)

# character vector 
char_data = c("Will", "math", "test")
class(char_data)

# logical vector 
logi = c(T, F, T, T)
class(logi)

# factor 
fact_data = as.factor(c("Good", "Best", "Worst", "Neutral"))

```

1.  **Data Frames**: Data frames are essential for working with tabular data. Understand how to create, subset, and manipulate data frames.

```{r}
num_col = 5
num_row = 5 
data = matrix(runif(num_row * num_col), num_row, num_col)
test_df = data.frame(data)
names(test_df) = c('d1', 'd2', 'd3', 'd4', 'd5')

# OR 
test_2 = data.frame(
  d1 = c(5,4,3), 
  d2 = c(9,1,2)
)

test_2
```

1.  **Functions**: Know how to create and use functions in R.

2.  **Control Structures**: Mastery of control structures like loops (for, while), if-else statements, and switch statements is important.

3.  **Data Types and Structures**: Understand other data structures like lists, matrices, and arrays.

4.  **Reading and Writing Data**: Ability to read data from files (e.g., CSV, Excel) and write data to files.

5.  **Basic Statistical Functions**: Familiarity with functions like mean, median, sum, sd, etc. Definitely review sampling and distributions with quartiles like qpois, rpois etc.

    ```{r}

    ```

6.  **Plotting**: Basic plotting functions such as plot(), hist(), and boxplot().

### Tidyverse Essentials:

1.  **Tibbles**: Tibbles are a modern version of data frames in the tidyverse. Know how to create and manipulate tibbles.
2.  **Piping (%\>%)**: Mastery of the pipe operator for chaining operations (%\>%).
3.  **dplyr**: Understand the dplyr package for data manipulation, including functions like filter(), select(), mutate(), summarise(), and arrange().
4.  **tidyr**: Know how to use tidyr functions like gather() and spread() for data tidying.
5.  **ggplot2**: Familiarity with ggplot2 for data visualization. Know how to create basic plots and customize them.
6.  **Stringr**: If dealing with text data, understanding functions in stringr for string manipulation.
7.  **Factors and Dates**: Tidyverse has its own way of handling factors and dates. Familiarize yourself with these methods if you'll be working with such data types.
8.  **Tidy Data Principles**: Understand the concept of tidy data and how to reshape data into tidy format using tidyr functions.
9.  **Purrr**: Know how to use purrr for functional programming, especially when working with lists and applying functions iteratively.

Remember to practice coding problems using both base R and tidyverse approaches to become proficient. Good luck with your interview preparation!
